Universal Anonymous Signatures

### 2022-05-12 prism-backend/research/gsac/app-gsac.tex
First complete readable draft. Still has some pending stuff and needs refinement, but should be shareable.

\todo{Need to update the proofs with the updated definitions, now returning
  \attrs as part of \Open.}

\commentwho{Jesus}{Strictly, in our generic construction we require IND-CCA, but
  in the following concrete instantiation, we use ElGamal, which is IND-CPA.
  However, we actually ``lift'' ElGamal with NIZKs of the encrypted plaintexts,
  which is essentially IND-CCA as far as I know. Still, make sure to mention
  that.}

\begin{proof}[\thmref{thm:correctness-gsac}; Correctness of \GSACGen]
  \todo{XXX}
\end{proof}

@@

\paragraph{$\OKeyGen(\parm) \rightarrow (\opk,\osk)$.} %
Outputs $\osk \gets \ZZ^*_p$, and $\opk \gets \gen{g}^{\osk}$. This is the
opener's ElGamal encryption key pair \needcite.

\paragraph{$\UKeyGen(\parm) \rightarrow (\upk,\usk)$.} %
Outputs $\usk \gets \ZZ^*_p$, and $\upk \gets \gen{h}_1^{\usk}$. \upk will
simply be used to compute Pedersen commitments \needcite to \usk. However,
it is useful to precompute it, and we treat it as a sort of ``public key''.

>>>>

\jdv{Strictly, in our generic construction we require IND-CCA, but
  in the following concrete instantiation, we use ElGamal, which is IND-CPA.
  However, we actually ``lift'' ElGamal with NIZKs of the encrypted plaintexts,
  but this not may be enough. Check and look for alternatives if not good.}

\begin{proof}[\thmref{thm:correctness-gsac}; Correctness of \GSACGen]
  By correctness of \SBCM and the NIZK for $\NIZKRel^{\Issue}$, an honestly
  obtained credential \cid, from an honest issuer, is a valid \SBCM signature
  over
  attribute set $\ATTR[\cid]$, associated to the user key of some honest user
  which, by convention, we denote with $\OWNR[\cid]$ -- thus, the secret key
  bound to the credential is $\PRVUK[\OWNR[\cid]]=\usk$. Since, at line 3 of the
  \ExpGSACCorrect game the environment checks that the \dattrs set output by
  $\adv$ is contained within the chosen credential \cid, again by correctness of
  \SBCM, the encryption scheme, and correctness of the NIZK for
  $\NIZKRel_{\Sign}$, \Sign succeeds with overwhelming
  probability, and \sig contains an encryption, under \opk, of $\ATTR[\cid]$,
  and due to correctness of the commitment scheme \C, also of $\upk =
  \CCommit(\usk;0)$. Additionally, correctness of the NIZK for
  $\NIZKRel_{\Open}$ ensures that \Open is able to decrypt them back into \upk,
  and that $\attrs=
  \ATTR[\cid]$, respectively. Finally, correctness of the NIZK for
  $\NIZKRel_{\Sign}$ and the NIZK for $\NIZKRel_{\Open}$ ensure that \Verify and
  \Judge accept the signature and opening proof, respectively, with overwhelming
  probability.
\end{proof}

@@

\paragraph{$\OKeyGen(\parm) \rightarrow (\opk,\osk)$.} %
Outputs $\osk \gets \ZZ^*_p$, and $\opk \gets \gen{g}^{\osk}$. This is the
opener's ElGamal encryption key pair.

\paragraph{$\UKeyGen(\parm) \rightarrow (\upk,\usk)$.} %
Outputs $\usk \gets \ZZ^*_p$, and $\upk \gets \gen{h}_1^{\usk}$. \upk will
simply be used to compute Pedersen commitments to \usk. However,
it is useful to precompute it, and we treat it as a sort of ``public key''.

### 2022-05-12 prism-backend/research/gsac/extensions-gsac.tex
Added high-level descriptions of how to build other schemes from UAS.
First complete readable draft. Still has some pending stuff and needs refinement, but should be shareable.

The approach to turn our non-interactive singing-verification into an
interactive protocol that ensures freshness is evident: prior to signing the
message, we require that the honest verifier sends to the user a number picked
uniformly at random from an appropriate domain, which must be concatenated to
the signed message. We sketch a proof for why does this result in a secure
\GSAC scheme with interactive signing and verification in
\appref{ssap:interactive-gsac}

@@

\todo{Can we do formal proofs here of $GSAC \implies$ GS and $GSAC \implies AC$?}

>>>>

The approach to turn our non-interactive singing-verification into an
interactive protocol that ensures freshness is evident: prior to signing the
message, we require that the honest verifier sends to the user a number picked
uniformly at random from an appropriate domain, which must be concatenated to
the signed message. If there is no need to actually sign a message, then the
message is set to an empty string, and the user only signs the random number.
%
Syntactically, instead of having non-interactive $\sig \gets \Sign(\gpk,\usk,
\cred,\dattrs,\msg)$ and $1/0 \gets \Verify(\gpk,\sig,\dattrs,\msg)$ algorithms,
we have a $1/0 \gets \langle \Sign(\gpk,\usk,\cred,\dattrs),\Verify(\gpk,
\dattrs) \rangle$ interactive protocol.
%
We sketch a proof for why does this result in a secure
\GSAC scheme with interactive signing and verification in
\appref{ssap:interactive-gsac}

@@

\jdv{Can we do formal proofs here of $GSAC \implies$ GS and $GSAC \implies AC$?}

### 2022-05-12 prism-backend/research/gsac/introduction.tex
First complete readable draft. Still has some pending stuff and needs refinement, but should be shareable.

% What is the context for this research? Introduce previous works on this topic,
% in one paragraph. (Privacy-preserving identities, with accountability (APPI))
Successful identity representation, management, and usage are, on its own, among
the toughest problems in the Internet and digital communications at large
\needcite. Just as an illustration, think
of usernames and passwords, which end up being unmanageable as soon as we need
to start remembering a few \needcite; or traditional vertical PKIs, like X.509,
which are too rigid, and frequently require too much trust \needcite; or
horizontal PKIs, such as PGP, that are also too rigid, and do not scale well
\needcite. If, in addition, we require privacy, the scenario becomes even more
challenging. Technologies like OpenID and OAuth \needcite make identity more
manageable (from the point of view of service providers) through Single Sign
On solutions. But this comes at the cost of introducing centralization points
that gain way too much knowledge beyond what is actually necessary. Holding
users accountable, or extracting utility out of their actions, is trivial in
this cases -- although that comes at the cost of barely having privacy.

@@

Concerning (1), either the signer is anonymous (before opening); or fully
deanonymized (after opening). Some variants on the traditional opening do not
fully deanonymize, but incorporate some other strict tradeoff. Like giving the
users full control \cite{dl21}, only allowing linking \needcite, or some
arbitrary extension such as allowing deanonymization only to the eyes of other
group members \needcite, or sequential linkability \cite{dl21}. While these
variants may be very suitable for concrete settings, it is not hard to see that
adapting them to other scenarios will be hard -- if possible at all.

@@

In fact, taking this unifying path is a low hanging fruit if one looks at recent
works. For instance, BBS+ signatures \cite{asm06,cdl16b} have been used recently
to build both group signatures \cite{gl19,dl21,fgl21}, but are also one of the
main candidates to incorporate anonymity into verifiable credentials%
\footnote{``W3C Credential Exchange using BBS+ Signatures'',
  \url{https://github.com/hyperledger/aries-rfcs/blob/main/features/0646-bbs-credentials/README.md}. Last access May 11th, 2020.}.%
The interest of the verifiable credentials community in BBS+ is not accidental,
though, as BBS+ follow the CL signatures line of work \cite{cl02}. CL signatures
allow creating a single signature over a set of messages, which can be directly
used to bundle a set of attributes (messages) into the same signature
(credential). Indeed, this was the trigger for some of the first, and possibly
the most fruitful branch of, anonymous credentials systems \cite{cl01,cl04}
\needcite. Other very closely related signature scheme, PS16 \cite{ps16} can
also be used to build group signatures, see \cite[Appendix A.1]{ps15} or
\cite{cdl+20}; as well as anonymous credentials \cite[Appendix A.2]{ps15} or
\cite{sms+19}.

@@

Additionally, within the AC domain, some works allow proving arbitrary claims
over the attributes in a credential \needcite; while others simply allow
selectively disclose subsets of these attributes. For \GSAC, we do the latter.
Despite not being the most general option (although it is more general than the
GS approach, as it does not even have attributes), it is enough to illustrate
the advantages of unifying both worlds. Again, in \UAS we generalize this
aspect.

Finally, modelling-wise we also have two choices: follow a UC-based definition
\needcite, or a game-based one \needcite. Independently on the pros and cons of
each, in GS the standard is the game-based approach
\cite{bmw03,bsz05,bcc+16,kty04,ky05}, although close relatives to GS, like DAA, prefer the
UC approach \needcite. On the AC line of work, while some UC models exist
\needcite, game-based definitions also seem prevalent \cite{fhs19} \needcite.
Thus, the logical choice (at least, for a first step) seems to be game-based
definitions. Concretely, we take strong inspiration from \cite{bsz05} for the
GS branch, and from \cite{fhs19} for the AC one.

@@

\paragraph{Generalizing issuance protocols.} %
In \GSAC, it is also not possible to leverage previously obtained credentials in
order to obtain new ones -- from the same issuer, or from another. Note that, if
we want to be really flexible, addressing this does not just mean letting the
user specify a set of credentials in the interactive protocol for obtaining a
new one. Concretely: What do we want to allow the user to prove with those extra
credentials? This is essentially the same case as in interactive credential
presentations. Thus, it seems reasonable to allow the same flexibility: namely,
proving arbitrary claims over the attributes in the credentials involved in the
request, without revealing them. We make only one exception, though: we do not
include support to blindly carry attributes from one credential used in the
request, to the new credential. This is indeed possible in some schemes
\needcite. However, we argue that, by including support to merge multiple
credentials into the same signature, this is no longer necessary. And indeed
it is even realistic, as it does not make much sense to expect that an issuer
attests an attribute that he cannot access in the clear (even despite it being
attested by some other issuer). In any case, adding support to prove arbitrary
claims at issuance time does introduce additional modelling needs. Namely, we
need to give a generic predicate definition, say \fissue. Then, our model needs
to make sure that, during the interactive issuance protocol, satisfaction of
this predicate is ensured. This is not straightforward, though. One option is,
as \cite{ckl+15} does for their issuance tokens, interpret that, if no
illegitimately obtained credential is used to produce a signature, then this
does no harm; so we only need to check at signature verification time that all
involved credentials were legitimately obtained. However, this turns into a
convoluted definition, which requires to mix claims of actions performed during
credential issuance with actions performed during signing (credential
presentation in \cite{ckl+15}). Instead, we opt for creating two separate
unforgeability-related properties: one capturing security at issuance time, and
another capturing security at signature creation time. Naturally, both require
honest issuers, as we detail later on.

@@

\paragraph{Universality of \UAS.} %
As stated, a main goal of \UAS is to reach a general enough model that covers a
large enough portion of the functionality modelled by previous GS and AC works,
and even goes beyond (e.g., through the \fissue, \feval and \finsp functions).
To showcase the universality of \UAS, we describe how, given a generic
construction of \UAS, the \fissue, \feval, and \finsp functions can be used to
come up with generic constructions of very well known schemes in the GS and AC
domains. Directly, \UAS is a generalization of \GSAC, which we prove to be a
strict generalization of vanilla GS and vanilla AC. Consequently, vanilla GS and
AC are a strict subset of \UAS. We also show how to build MPS schemes \needcite
which is on itself a generalization of BiAS \cite{lnpy21}; as well as group
signatures with message-dependent opening \cite{ehk+19}, or delegatable
anonymous credentials \needcite.
%
Interestingly, ``pure'' ring signatures \cite{rst06} can also be built from
\UAS, where with ``pure'' we mean that \UAS-based ring signatures do not require
an actual issuer -- i.e., users can act as issuers. This may even be suprising,
given that we build on
group signatures, where having no issuer of group credentials is just not
possible\footnote{Nobatly, mimicking ring signatures without issuers is
  trivially not possible in MPS, as its model does not support multiple
  issuers. This seems crucial as, in essence, a ring signature is a group
  signature where every user can become an issuer.}. Moreover, the
unforgeability notion, that requires trust on the issuer in the setting of group
signatures, is still applicable here, as long as the verifier trusts all
the public keys involved in the \UAS-based ring signature (just the same
as in conventional ring signatures). \todo{Although, on this regard, it is
  interesting to explore how attributes come into the picture.}

>>>>

% What is the context for this research? Introduce previous works on this topic,
% in one paragraph. (Privacy-preserving identities, with accountability (APPI))
Successful identity representation, management, and usage are, on its own, among
the toughest problems in the Internet and digital communications at large%
\footnote{``Digital identity is one of the 21st century's biggest challenges'':
  \url{https://www.maddyness.com/uk/2021/11/04/digital-identity-is-one-of-the-21st-centurys-biggest-challenges/}. Last access, May 12th, 2022.}.
Just as an illustration, think
of usernames and passwords, which end up being unmanageable as soon as we need
to start remembering a few%
\footnote{``The problem of passwords and how to deal with it'': \url{https://www.computerweekly.com/feature/The-problem-of-passwords-and-how-to-deal-with-it}. Last access,
May 12th, 2022.}; or traditional vertical PKIs, like X.509 \cite{rfc5280},
which are too rigid, and frequently require too much trust; or
horizontal PKIs, such as PGP, that are also too rigid, and are hard to manage
and scale\footnote{``What's the matter with PGP?'': \url{https://blog.cryptographyengineering.com/2014/08/13/whats-matter-with-pgp/}. Last access, May 12th, 2022}.
If, in addition, we require privacy, the scenario becomes even more
challenging. Technologies like OpenID\footnote{\url{https://openid.net/}. Last
  access, May 12th, 2022.} and OAuth\footnote{\url{https://oauth.net/}. Last
  access, May 12th, 2022.} make identity more
manageable (from the point of view of service providers) through Single Sign
On solutions. But this comes at the cost of introducing centralization points
that gain way too much knowledge beyond what is actually necessary. Holding
users accountable, or extracting utility out of their actions, is trivial in
this cases -- although that comes at the cost of barely having privacy.

@@

Concerning (1), either the signer is anonymous (before opening); or fully
deanonymized (after opening). Some variants on the traditional opening do not
fully deanonymize, but incorporate some other strict tradeoff. Like giving the
users full control \cite{dl21}, only allowing de-anonymization based on the
signed message \cite{ehk+19}, or some
arbitrary extension such as allowing deanonymization only to the eyes of other
group members \cite{manu06}, or sequential linkability \cite{dl21}. While these
variants may be very suitable for concrete settings, it is not hard to see that
adapting them to other scenarios will be hard -- if possible at all.

@@

In fact, taking this unifying path is a low hanging fruit if one looks at recent
works. For instance, BBS+ signatures \cite{asm06,cdl16b} have been used recently
to build both group signatures \cite{gl19,dl21,fgl21}, but are also one of the
main candidates to incorporate anonymity into verifiable credentials%
\footnote{``W3C Credential Exchange using BBS+ Signatures'',
  \url{https://github.com/hyperledger/aries-rfcs/blob/main/features/0646-bbs-credentials/README.md}. Last access May 11th, 2020.}.%
The interest of the verifiable credentials community in BBS+ is not accidental,
though, as BBS+ follow the CL signatures line of work \cite{cl02}. CL signatures
allow creating a single signature over a set of messages, which can be directly
used to bundle a set of attributes (messages) into the same signature
(credential). Indeed, this was the trigger for some of the first, and possibly
the most fruitful branch of, anonymous credentials systems (see
\cite{cl01,cl02,cl04}, among others). Other very closely related signature
scheme, PS16 \cite{ps16} can
also be used to build group signatures, see \cite[Appendix A.1]{ps15} or
\cite{cdl+20}; as well as anonymous credentials \cite[Appendix A.2]{ps15} or
\cite{sms+19}.

@@

Additionally, within the AC domain, some works allow proving arbitrary claims
over the attributes in a credential \cite{bcc+09,cklm14,dmm+18,fhs19}; while
others simply allow to
selectively disclose subsets of these attributes. For \GSAC, we do the latter.
Despite not being the most general option (although it is more general than the
GS approach, as it does not even have attributes), it is enough to illustrate
the advantages of unifying both worlds. Again, in \UAS we generalize this
aspect.

Finally, modelling-wise we also have two choices: follow a UC-based definition
\cite{cane20}, or a game-based one \cite{shoup04}. Independently on the pros and
cons of each, in GS the standard is the game-based approach
\cite{bmw03,bsz05,bcc+16,kty04,ky05}, although close relatives to GS, like DAA,
prefer the UC approach \cite{cdl16,cdl16b,ccd+17,cdl17}. On the AC line of work,
while some UC models exist \cite{}, game-based definitions also seem prevalent
\cite{bcc+09,cks10,cl11,cklm12,cmz14,dmm+18,fhs19}.
Thus, the logical choice (at least, for a first step) seems to be game-based
definitions. Concretely, we take strong inspiration from \cite{bsz05} for the
GS branch, and from \cite{fhs19} for the AC one.

@@

\paragraph{Generalizing issuance protocols.} %
In \GSAC, it is also not possible to leverage previously obtained credentials in
order to obtain new ones -- from the same issuer, or from another. Note that, if
we want to be really flexible, addressing this does not just mean letting the
user specify a set of credentials in the interactive protocol for obtaining a
new one. Concretely: What do we want to allow the user to prove with those extra
credentials? This is essentially the same case as in interactive credential
presentations. Thus, it seems reasonable to allow the same flexibility: namely,
proving arbitrary claims over the attributes in the credentials involved in the
request, without revealing them. We make only one exception, though: we do not
include support to blindly carry attributes from one credential used in the
request, to the new credential. This is indeed possible in some schemes
\cite{bcl04}. However, we argue that, by including support to merge multiple
credentials into the same signature, this is no longer necessary. And indeed
it is even realistic, as it does not make much sense to expect that an issuer
attests an attribute that he cannot access in the clear (even despite it being
attested by some other issuer). In any case, adding support to prove arbitrary
claims at issuance time does introduce additional modelling needs. Namely, we
need to give a generic predicate definition, say \fissue. Then, our model needs
to make sure that, during the interactive issuance protocol, satisfaction of
this predicate is ensured. This is not straightforward, though. One option is,
as \cite{ckl+15} does for their issuance tokens, interpret that, if no
illegitimately obtained credential is used to produce a signature, then this
does no harm; so we only need to check at signature verification time that all
involved credentials were legitimately obtained. However, this turns into a
convoluted definition, which requires to mix claims of actions performed during
credential issuance with actions performed during signing (credential
presentation in \cite{ckl+15}). Instead, we opt for creating two separate
unforgeability-related properties: one capturing security at issuance time, and
another capturing security at signature creation time. Naturally, both require
honest issuers, as we detail later on.

@@

\paragraph{Universality of \UAS.} %
As stated, a main goal of \UAS is to reach a general enough model that covers a
large enough portion of the functionality modelled by previous GS and AC works,
and even goes beyond (e.g., through the \fissue, \feval and \finsp functions).
To showcase the universality of \UAS, we describe how, given a generic
construction of \UAS, the \fissue, \feval, and \finsp functions can be used to
come up with generic constructions of very well known schemes in the GS and AC
domains. Directly, \UAS is a generalization of \GSAC, which we prove to be a
strict generalization of vanilla GS and vanilla AC. Consequently, vanilla GS and
AC are a strict subset of \UAS. We also show how to build MPS schemes \needcite
which is on itself a generalization of BiAS \cite{lnpy21}; as well as group
signatures with message-dependent opening \cite{ehk+19}, or delegatable
anonymous credentials \cite{bcc+09,cklm14}.
%
Interestingly, ``pure'' ring signatures \cite{rst06} can also be built from
\UAS, where with ``pure'' we mean that \UAS-based ring signatures do not require
an actual issuer -- i.e., users can act as issuers. This may even be suprising,
given that we build on
group signatures, where having no issuer of group credentials is just not
possible\footnote{Nobatly, mimicking ring signatures without issuers is
  trivially not possible in MPS, as its model does not support multiple
  issuers. This seems crucial as, in essence, a ring signature is a group
  signature where every user can become an issuer.}. Moreover, the
unforgeability notion, that requires trust on the issuer in the setting of group
signatures, is still applicable here, as long as the verifier trusts all
the public keys involved in the \UAS-based ring signature (just the same
as in conventional ring signatures). \jdv{Although, on this regard, it is
  interesting to explore how attributes come into the picture.}
